---
title: "Robustness Checks - Rosenfield et al. 2018"
author: "Kevin A. Rosenfield"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Read in data


#### Data (in .csv format) is read into a dataframe, look-times are converted from frames (29/second) to seconds, and descriptive statistics for looking behavior are displayed.


```{r tidy=TRUE}

Rhesus.Trials <- as.data.frame(read.csv("Experimental_Trial_Data.csv"))

Rhesus.Trials$Look_time <- Rhesus.Trials$Look_time / 29

Rhesus.Trials$Masc_time <- Rhesus.Trials$Masc_time / 29

Rhesus.Trials$Fem_time <- Rhesus.Trials$Fem_time / 29

```

```{r subset}
High_diff <- subset(Rhesus.Trials, MD_cat2 > 0)
Low_diff <- subset(Rhesus.Trials, MD_cat2 < 1)
```


#### Summaries for the two resulting datasets:


```{r summary high}

summary(High_diff[,c(5,24,30,31,32)])

```


```{r summary low}

summary(Low_diff[,c(5,24,30,31,32)])

```


#### Now, Wilcoxon signed-rank tests (paired, non-parametric) are repeated, once for the low masculinity differences group, once for the high masculinity differences group:


```{r prediction2.2a - high differences}

Prediction2.2a <- wilcox.test(High_diff$Masc_time, High_diff$Fem_time, paired = TRUE);

Prediction2.2a

Zstat2.2a<-qnorm(Prediction2.2a$p.value/2)
Effect_Size2.2a <- abs(Zstat2.2a)/sqrt(20);

Effect_Size2.2a

```


```{r prediction2.2b - low differences}

Prediction2.2b <- wilcox.test(Low_diff$Masc_time, Low_diff$Fem_time, paired = TRUE);

Prediction2.2b

Zstat2.2b<-qnorm(Prediction2.2b$p.value/2)
Effect_Size2.2b <- abs(Zstat2.2b)/sqrt(20);

Effect_Size2.2b

```