---
title: "Main Analyses - Rosenfield et al. 2018"
author: "Kevin A. Rosenfield"
date: 25 August 2018
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Install and load ggplot2

```{r ggplot2}
library(ggplot2)
```


## Read in trial data


* Data (in .csv format) is read into a dataframe, look-times are converted from frames (29/second) to seconds, and descriptive statistics for looking behavior are displayed.


```{r read_in}
Rhesus.Trials <- as.data.frame(read.csv("Experimental_Trial_Data.csv"))

Rhesus.Trials$Look_time <- Rhesus.Trials$Look_time / 29

Rhesus.Trials$Masc_time <- Rhesus.Trials$Masc_time / 29

Rhesus.Trials$Fem_time <- Rhesus.Trials$Fem_time / 29

summary(Rhesus.Trials[,c(5,24,30,31,32)])
```



### Did subjects look significantly longer at one stimulus type than the other?

* Boxplots depicting look-times for masculine and feminine stimuli

```{r boxplots1}

Rhesus_Trials2 <- Rhesus.Trials[,c(31,32)]

library(reshape) 
Rhesus_Trials2 <- melt(Rhesus_Trials2)

Rhesus_Trials2$Stim <- Rhesus_Trials2[,1]
Rhesus_Trials2$Time <- Rhesus_Trials2[,2]
Rhesus_Trials2 <- Rhesus_Trials2[,c(3,4)]

ggplot(Rhesus_Trials2, aes(x=Stim, y=Time, fill=Stim)) + ylab("Look-time (seconds)") + xlab("Image type")+ geom_boxplot() + coord_cartesian(ylim=c(0,11.2)) + scale_y_continuous(breaks=c(2, 4, 6, 8, 10)) + geom_jitter(shape=16, position=position_jitter(.2)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.position = "none") + theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold")) + geom_text(aes(1.5, 10.5, label="NS"), size=5)

```
  

* The following analysis includes the entire dataset (107 trials). Wilcoxon signed-rank test (paired, non-parametric):


```{r prediction1.1}

Prediction1.1 <- wilcox.test(Rhesus.Trials$Masc_time, Rhesus.Trials$Fem_time, paired = TRUE)

Prediction1.1

```


* We find no significant difference in masculine and feminine stimulus look-times.


### Did more subjects look longer at one stimulus type than the other?


* The following analysis includes the entire dataset (107 trials). Binomial test: is 64/105 different from chance?; 2 ties were automatically excluded:


```{r prediction 1.2}

Prediction1.2 <- binom.test(64, 105, p = 0.5,
           alternative = c("two.sided"),
           conf.level = 0.95);

Prediction1.2

```

* Our subjects looked longer at masculine stimuli than feminine stimuli in a significantly higher proportion of trials than would be predicted by chance.


### Was the proportion of time subjects spent looking at masculine images associated with the relative difference in masculinity between the two stimuli?

* Scatterplot depicting the relationship between masculinity difference scores and masculinity look-time percentages. Shaded area is th 95% confidence interval of the line.

```{r prediction 2.1a plot}

ggplot(Rhesus.Trials, aes(x=Masc_perc, y=Masc_diff, fill=Masc_perc)) + ylab("Masculinity difference score") + xlab("Masculine image look-time\n(as proportion of total look time)") + geom_smooth(method = 'lm') + geom_point(data=NULL, cex=1.5) + theme(panel.border = element_blank(), axis.line = element_line(colour = "black")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.position="none") + xlim(0,100) + theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold"))

```

1. The following analysis includes the entire dataset (107 trials). Linear model: predictor variable: Masculinity difference; dependent variable: masculine stilmulus look percentage:


```{r prediction 2.1a}

Prediction2.1a <- lm(Masc_perc ~ Masc_diff, data = Rhesus.Trials); 

summary(Prediction2.1a)

```

* The within-pair Masculinity difference explains a significant proportion of the variability in the masculine stilmulus look percentage.

2. Now we repeat the model, adding raw masculinity score as an additional predictor variable:

```{r prediction 2.1b}

Prediction2.1b <- lm(Masc_perc ~ Masc_diff + Masc_score, data = Rhesus.Trials); 

summary(Prediction2.1b)

```

* The within-pair Masculinity difference explains a significant proportion of the variability in the masculine stilmulus look percentage when controlling for the raw score of the masculinie image.

* The masculine image's score, however, does not explain a significant proportion of the variability in the masculine stilmulus look percentage when relative scores are controlled.


### Did subjects look significantly longer at one stimulus type than the other in  a subset of trials with high relative masculinity differences?


1. First, the dataset is subsetted into two data frames, based on a binary variable indicating whether the masculinity difference is high (n=54) or low (n=53).

* Boxplots depicting look-times for masculine and feminine stimuli in the high and low masculinity differences groups (requires changing data from horizontally to vertically oriented)

```{r boxplots2}

Rhesus_Trials2 <- Rhesus.Trials[,c(21, 31,32)]

Rhesus_Trials2_high <- subset(Rhesus_Trials2, MD_cat == 1)
Rhesus_Trials2_low <- subset(Rhesus_Trials2, MD_cat == 0)
Rhesus_Trials2_high <- Rhesus_Trials2_high[,c(2,3)]
Rhesus_Trials2_low <- Rhesus_Trials2_low[,c(2,3)] 

Rhesus_Trials2_high <- melt(Rhesus_Trials2_high)
Rhesus_Trials2_low <- melt(Rhesus_Trials2_low)

Rhesus_Trials2_high$Stim <- Rhesus_Trials2_high[,1]
Rhesus_Trials2_high$Time <- Rhesus_Trials2_high[,2]
Rhesus_Trials2_high <- Rhesus_Trials2_high[,c(3,4)]

Rhesus_Trials2_low$Stim <- Rhesus_Trials2_low[,1]
Rhesus_Trials2_low$Time <- Rhesus_Trials2_low[,2]
Rhesus_Trials2_low <- Rhesus_Trials2_low[,c(3,4)]


ggplot(Rhesus_Trials2_high, aes(x=Stim, y=Time, fill=Stim)) + ylab("Look-time (seconds)") + xlab("Image type")+ geom_boxplot() + coord_cartesian(ylim=c(0,11.2)) + scale_y_continuous(breaks=c(2, 4, 6, 8, 10)) + geom_jitter(shape=16, position=position_jitter(.2)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.position = "none") + theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold")) + geom_text(aes(1.5, 10.5, label="*"), size=10)

ggplot(Rhesus_Trials2_low, aes(x=Stim, y=Time, fill=Stim)) + ylab("Look-time (seconds)") + xlab("Image type") + geom_boxplot() + coord_cartesian(ylim=c(0,11.2)) + scale_y_continuous(breaks=c(2, 4, 6, 8, 10)) + geom_jitter(shape=16, position=position_jitter(.2)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + theme(legend.position = "none") + theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold")) + geom_text(aes(1.5, 10.5, label="NS"), size=5)

```

* Subsetting of data for analyses

```{r subset1}
High_diff <- subset(Rhesus.Trials, MD_cat == 1)
Low_diff <- subset(Rhesus.Trials, MD_cat == 0)
```


2. Summaries for the two resulting datasets:

+ High differences

```{r summary high differences}

summary(High_diff[,c(5,24,30,31,32)])

```

+ Low differences

```{r summary low differences}

summary(Low_diff[,c(5,24,30,31,32)])

```


3. Now, Wilcoxon signed-rank tests (paired, non-parametric) are repeated, once for the low masculinity differences group, once for the high masculinity differences group:

+ High differences

```{r prediction2.2a - high differences}

Prediction2.2a <- wilcox.test(High_diff$Masc_time, High_diff$Fem_time, paired = TRUE)

```

* We find that subjects looked significantly longer at masculine than feminine stimuli in the subset of trials with high masculinity differences.

+ Low differences

```{r prediction2.2b - low differences}

Prediction2.2b <- wilcox.test(Low_diff$Masc_time, Low_diff$Fem_time, paired = TRUE)

```

* We find no significant difference in masculine and feminine stimulus look-times in the subset of trials with low masculinity differences.


### Check for confounding variables

* Does group co-membership between subjects and stimulus males influence looking behavior?

1.

```{r subset2}

No_Groupmate <- subset(Rhesus.Trials, Groupmates=="None", select = "Masc_perc")

Fem_Groupmate <- subset(Rhesus.Trials, Groupmates=="Fem", select = "Masc_perc")

Masc_Groupmate <- subset(Rhesus.Trials, Groupmates=="Masc", select = "Masc_perc")

Both_Groupmate <- subset(Rhesus.Trials, Groupmates=="Both", select = "Masc_perc")

Groupmates <- merge(No_Groupmate, Fem_Groupmate)
```


2. Analyses (Wilcoxon signed-rank) testing whether subjects' masculine stimulus look-time percentages differed when the subject and 1) the feminine stimulus male was a groupmate, 2) the masculine stimulus male was a groupmate, or 3) both stimulus males were groupmates; each is compared to trials in which neither was a groupmate:


```{r groupmate}

confound_masc_groupmate <- wilcox.test(Masc_Groupmate$Masc_perc, No_Groupmate$Masc_perc, paired = FALSE); confound_masc_groupmate

counfound_fem_groupmate <- wilcox.test(Fem_Groupmate$Masc_perc, No_Groupmate$Masc_perc, paired = FALSE); counfound_fem_groupmate

confound_both_groupmxate <- wilcox.test(Both_Groupmate$Masc_perc, No_Groupmate$Masc_perc, paired = FALSE); confound_both_groupmxate

```


* We find no evidence that co-goup membership affected subjects'  looking behavior.

## Read in facial masculinity, color, and luminance data


```{r facial data}

Rhesus.Masculinity <- as.data.frame(read.csv("Facial_Masculinity_Data.csv"))

Rhesus.Masculinity$Color <- as.numeric(as.character(Rhesus.Masculinity$Color))

Rhesus.Masculinity$Luminance <- as.numeric(as.character(Rhesus.Masculinity$Luminance))

```

* Subset data into three datasets: One including all stumli, one including masculine stimuli only, and one including feminine stimuli only

```{r subset3}

Stimuli <- subset(Rhesus.Masculinity, M2_F1_Not0 > 0, select = c("Age","Color","Luminance","Masculinity_score"))

Masc_Stimuli <- subset(Rhesus.Masculinity, M2_F1_Not0 == 2, select = c("Age","Color","Luminance","Masculinity_score"))

Fem_Stimuli <- subset(Rhesus.Masculinity, M2_F1_Not0 == 1, select = c("Age","Color","Luminance","Masculinity_score"))

```

1.  Are facial masculinity scores related to age?

```{r age}

Age_corr <- cor.test(Stimuli$Masculinity_score, Stimuli$Age, method = "spearman");

Age_corr

Age_means <- wilcox.test(Masc_Stimuli$Age, Fem_Stimuli$Age, paired = FALSE);

Age_means

```

* Among stimuli, age is not signicantly correlated with facial masculinity, and age does not differ significantly between masculine and feminine stimuli.

2. Are facial masculinity scores related to facial color?

```{r color}

Color_corr <- cor.test(Stimuli$Masculinity_score, Stimuli$Color, method = "spearman");

Color_corr

Color_means <- wilcox.test(Masc_Stimuli$Color, Fem_Stimuli$Color, paired = FALSE);

Color_means

```

* Among stimuli, facial color is not signicantly correlated with facial masculinity, and facial color does not differ significantly between masculine and feminine stimuli.

3. Are facial masculinity scores related to facial luminance?

```{r luminance}


Luminance_corr <- cor.test(Stimuli$Masculinity_score, Stimuli$Luminance, method = "spearman");

Luminance_corr

Luminance_means <- wilcox.test(Masc_Stimuli$Luminance, Fem_Stimuli$Luminance, paired = FALSE); 

Luminance_means

```

*Among stimuli, facial luminance is not signicantly correlated with facial masculinity, and facial luminance does not differ significantly between masculine and feminine stimuli.

### Analyses of possible confounding variables reveals no evidence that facial masculinity is confounded by the age, facial color, or facial luminance of stimulus males.




#### Because we seperated the odd number of trials into high (N=54) and low (N=53) masculinity differences groups arbitrarily, to test robustness of our results, we repeated these analyses with trials seperated into alternative high (N=53) and low (N=54) groups (one trial moved the high to the low group).

```{r subset extra}
High_diff_extra <- subset(Rhesus.Trials, MD_cat2 > 0)
Low_diff_extra <- subset(Rhesus.Trials, MD_cat2 < 1)
```


#### Summaries for the two resulting datasets:


```{r summary high extra}

summary(High_diff_extra[,c(5,24,30,31,32)])

```


```{r summary low extra}

summary(Low_diff_extra[,c(5,24,30,31,32)])

```


#### Now, Wilcoxon signed-rank tests (paired, non-parametric) are repeated, once for the low masculinity differences group, once for the high masculinity differences group:


```{r prediction2.2a - high differences extra}

Prediction2.2a_extra <- wilcox.test(High_diff$Masc_time, High_diff$Fem_time, paired = TRUE);

Prediction2.2a_extra

```

* We still find that subjects looked significantly longer at masculine than feminine stimuli in the subset of trials with high masculinity differences.

```{r prediction2.2b - low differences extra}

Prediction2.2b_extra <- wilcox.test(Low_diff$Masc_time, Low_diff$Fem_time, paired = TRUE);

Prediction2.2b_extra

```

* We still find no significant difference in masculine and feminine stimulus look-times in the subset of trials with low masculinity differences.

